---
title: "Evaluating the Influence of Marine Protected Areas on Surf Zone Fish"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r libraries, echo = FALSE, warning = FALSE, message = FALSE}
library("lme4")
library("readxl")
library("ggh4x")
library("dplyr")
library("AICcmodavg")
library("MuMIn")
require("pscl")
require("MASS")
require("boot")
library("DHARMa")
library("emmeans")
library("ggthemes")

library("lmerTest")

suppressPackageStartupMessages(library("tidyverse")) 
suppressPackageStartupMessages(library("janitor"))
suppressPackageStartupMessages(library("knitr")) # to kable
suppressPackageStartupMessages(library("flextable")) #to flextable

# set seed so analyses are reproducible 
set.seed(123)
```

Our goal is to investigate the ecological effectiveness of California MPAs for surfzone fish - through biological metrics of richness, abundance, and biomass. To do this we used MPA status and region as fixed effects in a mixed effect model. Random effects for site pairs were included to account possible spatial autocorrelation between nearby sites and to account for any variability not accounted for in the region term (Pinheiro and Bates 2000). A log(x+1) transformation for all response variables. Multicollinearity between predictor variables, distributions of residuals, normality and over-dispersion were checked using variance inflation factors for mixed models, residual plots and Q-Q plots and were all found to be acceptable in the final mixed model. 

The following code illustrates the workflow employed in this manuscript and acts as an example of they analyses conducted. Additional analyses are described in the text and can be recreated from the data in this workflow.

## Seines
```{r load seine data, echo = FALSE, warning = FALSE, message = FALSE}
# seine observations
seines <- read.csv("../data/seine_fish_19_20.csv", row.names=1) 

```


```{r generate response metrics, echo = FALSE, warning = FALSE, message = FALSE}
# generate response metrics: CPUE, RPUE, BPUE
# generates a df of the number of hauls pulled per sampling day at each site
site_samplingdate_no_haul <- seines %>%
  group_by(region, mpa_status, site_name, site_pair,  year, month, day, month_day) %>%
  dplyr::summarise( num_hauls = length(unique(month_day_haul_no))) 

# counts abundance, richness, and biomass for each haul 
catch_summary_by_site_haul <- seines %>%
  filter(species_code !="NOSP") %>% # remove observations that did not catch any fish
  group_by(region, mpa_status, site_name, site_pair, haul_number, year, month, day, month_day) %>%
  dplyr::summarise(abund = sum(count), 
                   sp_richness = length(unique(species_code)),
                   bio = sum(fish_weight, na.rm = TRUE)) # how much is caught in each haul

catch_summary_by_site_sampling <- 
  site_samplingdate_no_haul %>%
  left_join(catch_summary_by_site_haul) %>%
  group_by(region, mpa_status, site_name, site_pair,  year, month, day, month_day) %>% 
  dplyr::reframe( num_hauls = num_hauls, 
                    # abundance
                    abundance = sum(abund), # add up all the hauls on a given sampling date
                    CPUE = abundance/num_hauls, # all sites have different haul numbers
                    # richness
                    species_richness = sum(sp_richness),
                    RPUE=species_richness/num_hauls,
                    # biomass
                    biomass = sum(bio),
                    BPUE= biomass/num_hauls
  ) %>% distinct()

catch_summary_by_site_sampling[is.na(catch_summary_by_site_sampling)] <- 0 # fill in the sampling dates at sites where no fish were caught

# Create a column to identify sampling event
catch_summary_by_site_sampling <- catch_summary_by_site_sampling %>%
  group_by(region, site_pair, site_name, year) %>%
  mutate(sampling_event = match(month_day, unique(month_day))) 

 catch_summary_by_site_sampling$mpa_status <- as.factor(catch_summary_by_site_sampling$mpa_status)
  catch_summary_by_site_sampling$region <- as.factor(catch_summary_by_site_sampling$region)
```


Richness model
```{r}
mod_log_rpue <- lmer(log1p(RPUE)~0+mpa_status*region+(1|site_pair),  data=catch_summary_by_site_sampling)
as_flextable(mod_log_rpue) %>%   colformat_double(digits = 2)
```
Pseudo $R^2$ results
```{r}
rsq_rpue <- r.squaredGLMM(mod_log_rpue)
```


Anova table of results
```{r rpue, echo = FALSE, warning = FALSE, message = FALSE}
aov_rpue <- anova(mod_log_rpue, type="3", ddf="Kenward-Roger")
as_flextable(aov_rpue) %>%   colformat_double(digits = 2)
```

Assumptions
```{r assumpt rpue, echo = FALSE, warning = FALSE, message = FALSE}
res_rpue <- simulateResiduals(mod_log_rpue, plot = T)# check for over dispersion 
```

Post hoc results
```{r posthoc rpue, echo = FALSE, warning = FALSE, message = FALSE}
post_rpue_int <- emmeans(mod_log_rpue, list(pairwise ~ mpa_status | region), adjust = "tukey", type = "response")
post_rpue_int
post_rpue_region <- emmeans(mod_log_rpue, list(pairwise ~ region ), adjust = "tukey", type = "response")
post_rpue_region
post_rpue_mpa <- emmeans(mod_log_rpue, list(pairwise ~ mpa_status ), adjust = "tukey", type = "response")
post_rpue_mpa
```



## BRUVs
```{r load maxN data}
maxn_all <- read.csv("../data/maxn_all.csv", row.names = 1)

op_code_list <- maxn_all %>%
  dplyr::select(region, ca_mpa_name_short,site_name, mpa_status, 
                op_code,site_pair, year, month, day, bruv) %>%
  as.data.frame() %>%
  distinct()

maxn_summary <- maxn_all %>% 
  filter(class == "Actinopteri" | class == "Elasmobranchii") %>% 
   right_join(op_code_list) %>% 
  group_by(region, op_code, site_name,ca_mpa_name_short, site_pair, year, month, day, bruv, mpa_status) %>%
  #
  dplyr::summarise(maxn=sum(max_n))
maxn_summary$maxn[is.na(maxn_summary$maxn)] <- 0 #adds 0 value to bruvs with no fish present

mean_maxn <- maxn_summary %>% 
  group_by(region, site_name,ca_mpa_name_short, site_pair,  mpa_status, year, month, day) %>% 
  dplyr::summarise(
    no_bruv =length(unique(bruv)),
    maxn_survey = sum(maxn),
    mean = maxn_survey/no_bruv, # should be fine because they all have same number of BRUVs analyzed
    sd = sqrt(sum((maxn-mean)^2))/(no_bruv-1), # haul abundance - mean
    min = min(maxn), 
    max = max(maxn), 
    median = median(maxn), 
    n_maxn = n()) %>% 
  mutate(standard_error = (sd / sqrt(n_maxn))) %>% 
  arrange(region) 

mean_maxn$region <- as.factor(mean_maxn$region)
mean_maxn$mpa_status <- as.factor(mean_maxn$mpa_status)
```


```{r bruv richness data}
richness_summary <- maxn_all %>% 
  filter(class == "Actinopteri" | class == "Elasmobranchii") %>% 
  right_join(op_code_list) %>% 
  group_by(region, site_name,ca_mpa_name_short,site_pair, year, month, day, bruv, op_code,  mpa_status) %>% 
  dplyr::summarise(sp_rich=sum(taxonomic_unit)) 

richness_summary$sp_rich[is.na(richness_summary$sp_rich)] <- 0 #adds 0 value to bruvs with no fish present

mean_richness_sampling <- richness_summary %>% 
  group_by(region, site_name,ca_mpa_name_short, site_pair,  mpa_status,  year, month, day) %>% 
  dplyr::summarise(
    no_bruv =length(unique(op_code)),
    sp_rich_survey = sum(sp_rich),
    mean = sp_rich_survey/no_bruv) %>% 
  arrange(region)

mean_richness_sampling$region <- as.factor(mean_richness_sampling$region)
mean_richness_sampling$mpa_status <- as.factor(mean_richness_sampling$mpa_status)
```

BRUV richness model results
```{r bruv rich mod, echo = FALSE, warning = FALSE, message = FALSE}
mod_log_rich <- lmer(log1p(mean)~0+mpa_status*region+(1|site_pair),  data=mean_richness_sampling)
as_flextable(mod_log_rich) %>%   colformat_double(digits = 2)
```

Pseudo $R^2$
```{r}
rsq_rich <- r.squaredGLMM(mod_log_rich)
rsq_rich
```

Anova table results
```{r}
aov_rich <- anova(mod_log_rich, type="3", ddf="Kenward-Roger")
as_flextable(aov_rich) %>%   colformat_double(digits = 2)
```

Assumptions
```{r assumpt bruv rich, echo = FALSE, warning = FALSE, message = FALSE}
res_rich <- simulateResiduals(mod_log_rich, plot = T)# check for over dispersion NS!
```

Post hoc results
```{r posthoc bruv rich, echo = FALSE, warning = FALSE, message = FALSE}
post_rich_bruv_int <- emmeans(mod_log_rich, list(pairwise ~ mpa_status | region), adjust = "tukey", type = "response")
post_rich_bruv_int
post_rich_bruv_region <- emmeans(mod_log_rich, list(pairwise ~ region ), adjust = "tukey", type = "response")
post_rich_bruv_region
post_rich_bruv_mpa <- emmeans(mod_log_rich, list(pairwise ~ mpa_status ), adjust = "tukey", type = "response")
post_rich_bruv_mpa
```


## Elasmobranchs 
```{r elasmo maxn data, echo = FALSE, warning = FALSE, message = FALSE}
elasmo_sp_b <- maxn_all %>% filter(region =="South") %>% filter(class=="Elasmobranchii")
elasmo_sp_b$scientificname <- paste0(elasmo_sp_b$genus, " ", elasmo_sp_b$species)
#print(unique(elasmo_sp_b$scientificname))
maxn_summaryE <- elasmo_sp_b %>% 
  filter(class == "Elasmobranchii") %>% 
   right_join(op_code_list) %>% 
  group_by(region, op_code, site_name,ca_mpa_name_short, site_pair, year, month, day, bruv,  mpa_status) %>%
  dplyr::summarise(maxn=sum(max_n))

maxn_summaryE$maxn[is.na(maxn_summaryE$maxn)] <- 0 #adds 0 value to bruvs with no fish present

mean_maxnE <- maxn_summaryE %>% 
  group_by(region, site_name,ca_mpa_name_short, site_pair, mpa_status, year, month, day) %>% 
  dplyr::summarise(
    no_bruv =length(unique(bruv)),
    maxn_survey = sum(maxn),
    mean = maxn_survey/no_bruv) %>% 
  arrange(region) 
mean_maxnE$no_bruv <- mean_maxn$no_bruv
mean_maxnE <- mean_maxnE %>% 
  mutate(mean = maxn_survey/no_bruv)
mean_maxnE <- mean_maxnE %>% filter(region == "South") # the south did not watch 6 bruvs
```

Model summary
```{r maxn elasm mod}
mod_log_maxn_e <- lmer(log1p(mean)~0+mpa_status+(1|site_pair),  data=mean_maxnE)
as_flextable(mod_log_maxn_e) %>%   colformat_double(digits = 2)
```

Pseudo $R^2$
```{r}
rsq_maxn_elas <- r.squaredGLMM(mod_log_maxn_e) 
rsq_maxn_elas
```

Anova results table
```{r}
aov_maxn_elas <- anova(mod_log_maxn_e, type="3", ddf="Kenward-Roger")
as_flextable(aov_maxn_elas) %>%   colformat_double(digits = 2)
```

Assumptions
```{r assumpt maxn elasm, echo = FALSE, warning = FALSE, message = FALSE}
res_maxn_elas <- simulateResiduals(mod_log_maxn_e, plot = T)# check for over dispersion NS!
```

Post hoc 
```{r}
post_elas_maxn_mpa <- emmeans(mod_log_maxn_e, list(pairwise ~ mpa_status ), adjust = "tukey", type = "response")

ph_maxN_e  <- post_elas_maxn_mpa$`pairwise differences of mpa_status` %>%
     summary(infer = TRUE) %>%  data.frame()
```

## Figures

```{r rearrange data frames}
# make some new plots
library(ggh4x)
library('ggthemes')
# data
catch_site <- pivot_longer(catch_summary_by_site_sampling[, c(1:8,16,11,13,15)], 
                            cols = c("CPUE", "RPUE", "BPUE"),
                            names_to = "metric",
                            values_to = "value")

catch_site2 <-  catch_site %>% 
                dplyr::group_by(region, site_name, mpa_status, site_pair, metric) %>% 
                dplyr::summarize(median = median(value),
                          mean = mean(value),
                          qt1 = quantile(value, prob = 0.25),
                          qt3 = quantile(value, prob = 0.75),
                          CI_L = mean - (sd(value) * 1.96)/sqrt(length(unique(value))),
                          CI_H = mean + (sd(value) * 1.96)/sqrt(length(unique(value))))
catch_region <-  catch_site %>% 
  group_by(region, mpa_status, metric) %>% 
  dplyr::summarize(median = median(value),
            mean = mean(value),
            qt1 = quantile(value, prob = 0.25),
            qt3 = quantile(value, prob = 0.75),
            CI_L = mean - (sd(value) * 1.96)/sqrt(length(unique(value))),
            CI_H = mean + (sd(value) * 1.96)/sqrt(length(unique(value))))


catch_site2$region <- factor(catch_site2$region, levels = c("North", "Central", "South"))
catch_site2$metric <- factor(catch_site2$metric, levels = c("RPUE", "CPUE", "BPUE"))

catch_region$region <- factor(catch_region$region, levels = c("North", "Central", "South"))
catch_region$metric <- factor(catch_region$metric, levels = c("RPUE", "CPUE", "BPUE"))

# bruvs
maxn <- mean_maxn[, c("region", "site_name","site_pair" ,  "mpa_status", "year", "month", "day", "mean")]
maxn$metric <- "CPUE BRUV"
rich_bruvs <- mean_richness_sampling[, c("region", "site_name","site_pair" ,  "mpa_status", "year", "month", "day", "mean")]
rich_bruvs$metric <- "RPUE BRUV"
bruv <- rbind(maxn, rich_bruvs)
colnames(bruv)[colnames(bruv)=="mean"]<- "mu"

bruv_site <- bruv %>% 
  group_by(region, site_name, mpa_status, site_pair, metric) %>% 
  dplyr::summarize(median = median(mu),
            mean = mean(mu),
            qt1 = quantile(mu, prob = 0.25),
            qt3 = quantile(mu, prob = 0.75),
            CI_L = mean - (sd(mu) * 1.96)/sqrt(length(unique(mu))),
            CI_H = mean + (sd(mu) * 1.96)/sqrt(length(unique(mu)))
            )

seines_bruvs_site <- rbind(catch_site2, bruv_site)
seines_bruvs_site$region <- factor(seines_bruvs_site$region, levels = c("North", "Central", "South"))
seines_bruvs_site$metric <- factor(seines_bruvs_site$metric, levels = c("RPUE","RPUE BRUV", "CPUE", "CPUE BRUV","BPUE"))

bruv_region <- bruv %>% 
  group_by(region, mpa_status, metric) %>% 
  dplyr::summarize(median = median(mu),
            mean = mean(mu),
            qt1 = quantile(mu, prob = 0.25),
            qt3 = quantile(mu, prob = 0.75),
            CI_L = mean - (sd(mu) * 1.96)/sqrt(length(unique(mu))),
            CI_H = mean + (sd(mu) * 1.96)/sqrt(length(unique(mu))))

seines_bruvs_region <- rbind(catch_region, bruv_region)
seines_bruvs_region$region <- factor(seines_bruvs_region$region, levels = c("North", "Central", "South"))
seines_bruvs_region$metric <- factor(seines_bruvs_region$metric, levels = c("RPUE","RPUE BRUV", "CPUE", "CPUE BRUV","BPUE"))
```

```{r plot theme}

dodge <- position_dodge(width=0.5)

plot_theme <-  theme_tufte()+
  theme(axis.title=element_text(size=12, color="black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(vjust=0.95, size = 12),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, vjust = 0.95, hjust =0.95),
        axis.text.y = element_text(size = 12, color = "black"),
        axis.ticks = element_line(color="black"),
        axis.line = element_line(color="black"),
        panel.background=element_blank(),
        #panel.grid.major=element_blank(),
        legend.position="bottom",
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        strip.background=element_blank(),
        strip.placement = "outside",
        #strip.text=element_blank(),
        panel.spacing = unit(0.5, "cm","lines"))

```

```{r}
seines_bruvs_site %>%  ggplot(aes(x=site_pair, y=median, color= mpa_status))+
  geom_point(position = dodge, size = 2 )+
  geom_errorbar(aes(ymin = qt1, ymax = qt3), position = dodge, width = 0.2)+
  scale_color_manual(name = " ", values=c("#990000", "#3399FF")) + 
  facet_wrap2(metric ~ ., scales = "free_y",
             nrow = 5, 
             strip.position = "left",axes = "all", remove_labels = "x") + 
  labs(x= "Site Pair",
       y = " ")+ plot_theme
```

```{r}
seines_bruvs_region %>%  ggplot(aes(x=region, y=median, color= mpa_status))+
  geom_point(position = dodge, size = 3 )+
  geom_errorbar(aes(ymin = qt1, ymax = qt3), position = dodge, width = 0.2)+
  scale_color_manual(name = " ", values=c("#990000", "#3399FF")) + 
  facet_wrap2(metric ~ ., scales = "free_y",
              nrow = 5, 
              strip.position = "left", axes = "all", remove_labels = "all") + 
  labs(x= "Region",
       y = " ")+ plot_theme
```